---
title: "Scan Examples"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scan Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Isoprocessor supports several plotting and data conversion functions for scan data. This vignette shows some of the functionality for scan files read by the [isoreader package](http://isoprocessor.isoverse.org) (see the [corresponding vignette](http://isoreader.isoverse.org/articles/scan.html) for details on data retrieval, storage and export). 

```{r, warning=FALSE, message=FALSE}
# load isoreader and isoprocessor packages
library(isoreader)
library(isoprocessor)
```

# Reading files

```{r}
# read scan examples provided by the isoreader package
scan_files <- 
  iso_read_scan(
    iso_get_reader_example("peak_shape_scan_example.scn"),
    iso_get_reader_example("background_scan_example.scn"),
    iso_get_reader_example("full_scan_example.scn"),
    iso_get_reader_example("time_scan_example.scn")
  )
```

# Visualization

Plotting the raw data can be done either with the generic `iso_plot_raw_data()` function for a simple plot with default parameters, or directly using the scan specific `iso_plot_scan_data()`:

## Voltage Scan

```{r "all_data_HV", fig.width = 12, fig.asp = 0.8}
iso_plot_scan_data(scan_files, type = "High Voltage")
```

## Magnet Scan

```{r "all_data_magnet", fig.width = 12, fig.asp = 0.8}
iso_plot_scan_data(scan_files, type = "MagnetCurrent")
```

## Time Scan

```{r "all_data_clock", fig.width = 12, fig.asp = 0.8}
iso_plot_scan_data(scan_files, type = "Clock")
```

## Customization

All customization options are described in the function help (`?iso_plot_scan_data`) and include, for example, plotting only a specific subset of masses:

```{r "select_masses", fig.width = 10, fig.asp = 0.3}
# plot just masses 44 and 45
iso_plot_scan_data(
  scan_files, type = "High Voltage",
  data = c("44", "45")
)
```

# Isotope ratios

Isoprocessor provides a dynamic ratio calculation function (`iso_calculate_ratios()`) that accepts any combination of masses, here demonstrated for several different ratios. In this context, another useful customization option of the plotting function is the option to adjust plot aesthetics such as color, shape, and paneling:

```{r "ratios", fig.width = 8, fig.asp = 1}
scan_files <-
  scan_files %>% 
  # calculate 46/44, 45/44
  iso_calculate_ratios(ratios = c("46/44", "45/44"))

iso_plot_scan_data(
  scan_files, type = "High Voltage",
  # visualize masses and ratios
  data = c(44:46, "45/44", "46/44"),
  # focus on peak
  x_interval = c(9.44, 9.49),
  # panel including by category
  panel = category ~ file_id
)
```

# Signal conversion

Isoprocessor can convert between different signal units to examine the primary ion currents. 

```{r "signal_conversion", fig.width = 10, fig.asp = 0.3}
scan_files %>% 
  # convert all signals to nano ampere
  iso_convert_signals(to = "pA") %>% 
  # plot primary ion currents
  iso_plot_scan_data(data = c(44, 45))
```

# Plot styling

Since all isoprocessor plots are standard [ggplot](http://ggplot2.tidyverse.org/) objects, they can be modified with any ggplot commands. 

```{r "plot styling", message=FALSE, fig.width = 8, fig.asp = 0.7}
library(ggplot2)
# replot
iso_plot_scan_data(scan_files, type = "MagnetCurrent") +
  # modify plot styling (ggplot functionality)
  theme(text = element_text(size = 20))
```
